---
title: "aSPU for Meta Analized Data (aSPUM() and aSPUMpath())"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{aSPU for Meta Analized Data}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r knitr_options, echo=FALSE, results=FALSE}
library(knitr)
opts_chunk$set(fig.width = 12)
```

This vignette illustrates the use of the aSPU package for
meta-analyzed data.


## Data

We will consider the analysis of a coronary artery disease (CAD)
data from the CARDIoGRAM and C4D consortium. The data set contains P value
data for coronary artery disease (CAD). This study comprised 63,746
CAD cases and 130,681 controls. We mapped these SNPs to the 9th KEGG
pathway ([KEGG_STEROID_BIOSYNTHESIS](http://www.broadinstitute.org/gsea/msigdb/cards/KEGG_STEROID_BIOSYNTHESIS))
. Let's load this subset.

```{r loading, eval=FALSE}
library(aSPU)
data(kegg9)
```
The 9th Kegg pathway contains 16 genes.

```{r kegg1}
kegg9$gene.info
```

The `PPs` is a list object contains the SNP information for each genes.
```{r kegg2}
## SNPs mapped on 3rd and 4th genes of 9th Kegg pathway
kegg9$PPs[3:4]
```

The `Ps` object contains p-value information for mapped
SNPs. Total `r length(kegg9$Ps)`SNPs are mapped on 9th kegg pathway.

```{r kegg3}
length(kegg9$Ps)
kegg9$Ps[1:10]
```

Using [plink](http://pngu.mgh.harvard.edu/~purcell/plink/index.shtml),
we mapped our SNPs to the reference population ( Hapmap CEU phase 2
). We dropped the SNPs of minor allele frequency (MAF) less then 5
percent. Among `r length(kegg9$Ps)` SNPs mapped on 9th Kegg pathway, `r
length(kegg9$nP)` SNPs are mapped on the refference data. P-values of these
SNPs and correlation matirx of SNPs using the reference population,
saved on `nP` and `ldmatrix` object.

```{r kegg4}
kegg9$nP[1:10]
kegg9$ldmatrix[1:10,1:10]
```
The `snp.info` data object have snp information for each mapped
SNPs. The 1st column is rsID, 2nd column is Chr, 3rd column is
location and 4th column is p-value.

```{r kegg5}
kegg9$snp.info[1:10,]
```

## Use of aSPUM function

Using the following code we can use `aSPUM()` function to get gene-wise
p-value for each gene in 9th Kegg pathway.
```{r gene-wise p}
Gps<-NULL;
gl <- NULL;
for( g in kegg9$gene.info[,1]) {
    snps <-
    which( ( kegg9$snp.info[,2] == kegg9$gene.info[kegg9$gene.info[,1]==g,2]) &
           (kegg9$snp.info[,3] > kegg9$gene.info[kegg9$gene.info[,1] == g, 3])&
           (kegg9$snp.info[,3] < kegg9$gene.info[kegg9$gene.info[,1] == g, 4]))

        newP <- kegg9$nP[snps] ;
        ldsub <- kegg9$ldmatrix[snps, snps];

        if( length(snps) > 0) {
            out <- aSPUM(newP, corrSNP=ldsub , pow=c(1,2,4,8, Inf),
                         n.perm=1000, Ps=TRUE)
            Gps <- rbind(Gps, out$pvs)
            gl <- c(gl, g)
        }
    }
rownames(Gps) <- gl
Gps
```

## Use of aSPUMpath function

We can see get p-value for pathways as follows.
```{r pathway p}
out.g <- GatesSimes(pvec = kegg9$nP, ldmatrix = kegg9$ldmatrix,
                    snp.info=kegg9$snp.info, gene.info = kegg9$gene.info)

out.h <- Hyst(pvec = kegg9$nP, ldmatrix = kegg9$ldmatrix,
              snp.info=kegg9$snp.info, gene.info = kegg9$gene.info)

out.a <- aSPUMpath(kegg9$nP, corrSNP = kegg9$ldmatrix, pow=c(1:8, Inf),
                   pow2 = c(1,2,4,8),
                   snp.info=kegg9$snp.info, gene.info = kegg9$gene.info,
                   n.perm=1000, Ps = TRUE)

out.g; out.h; out.a

```
`out.g` is for Gates-Simes, `out.h` is for HYST and `out.a` is for
`aSPUMpath`.
